### 2016-12-20 paper/chapters/vae.tex
added

\section{Variational Autoencoders}
\label{sec:vae}
Variational Autoencoders (VAE) have been developed concurrently in 2014 by Kingma et al. \cite{vae:2014} and Rezende et al. \cite{dlgm:2014}.
The proposed framework includes in addition to an differentiable generator network an inference network in order to perform efficient inference.

Using an inference network allows the model to be trained fully using gradient-based methods and error backpropagation.

The relation to other autoencoders becomes appearant when the inference networks gets renamed to \emph{probabilistic encoder} $q_\theta(z|x)$ and the generator to \emph{probabilistic decoder} $p_\phi(x|z)$ whereas the encoder maps samples $x$ from the data space to latent space $z$, or more precisely parameters of the distribution $q_\theta(z|x)$.
Meanwhile does the decoder maps variables from the latent space $z$ back into the data space as shown in figure \ref{fig:vae_architecture}.\\

However it is not appearent how this model would provide sufficiently good results. In order to understand how well the variational autoencoder performs, it is important to understand the variational methods used to build the model.

@@

\subsection{Variational Inference}
The calculus of variations is used to perform optimization on a family of functions.
As such variational based methods have become a popular choice to deal with intractable posterior distributions occuring in inference problems.

>>>>

\section{Variational Autoencoders}
\label{sec:vae}
Variational Autoencoders (VAE) have been developed concurrently in 2014 by Kingma et al. \cite{vae:2014} and Rezende et al. \cite{dlgm:2014}.
The proposed framework includes in addition to a differentiable generator network an inference network with which the model is able to efficiently approximate the posterior distribution.
Using an inference network allows the model to be trained fully using gradient-based methods and error backpropagation.\\
%The relation to other autoencoders becomes appearant when the inference networks gets renamed to \emph{probabilistic encoder} $q_\theta(z|x)$ and the generator to \emph{probabilistic decoder} $p_\phi(x|z)$ whereas the encoder maps samples $x$ from the data space to latent space $z$, or more precisely parameters of the distribution $q_\theta(z|x)$.
%Meanwhile does the decoder maps variables from the latent space $z$ back into the data space as shown in figure \ref{fig:vae_architecture}.\\

However it is not appearent how this model would provide sufficiently good results. In order to understand how well the variational autoencoder performs, it is important to first understand the variational methods used to build the model.

@@

\paragraph{Inference problem}
Given a dataset $X = { x^{(1)}, x^{(2)}, \dots, x^{(n)}}$ and a model with observations $x \in X$ and hidden variables $z$. The inference problem is posed as computing the posterior distribution $p(z|x)$ 
\begin{equation}
  \label{eq:intractable_posterior}
  p(z|x) = \frac{p(x|z) p(z)}{p(x)}
\end{equation}
However, computing the right-hand side of equation \ref{eq:intractable_posterior} is intractable due to the exponentially many configurations of the hidden variables. As such it is often needed to use approximate inference such as variational inference.

\subsection{Variational Inference}
Variational inference (VI) can be used to turn the inference problem into an optmization problem which we are able to approximate using gradient-based methods.
Approximating a posterior distribution 
%The calculus of variations is used to perform optimization on a family of functions.
%As such variational based methods have become a popular choice to deal with intractable posterior distributions occuring in inference problems as seen in \ref{eq:intractable_posterior}.
